---
title: "LahmanSQL"
output: html_document
date: "2025-01-02"
---

## Introduction

The intention behind this exercise is to improve on my SQL abilities, and to dive into some baseball statistics. 

To begin, I downloaded the SQL version of Sean Lahman's iconic baseball database that spans the 1871-2023 seasons. You can find the database, and more about Sean Lahman here: http://seanlahman.com/. 


## Accessing the SQL database

The SQL database file was in a .bak format, which I was unfamiliar with initially. I learned that this was a common way to export a backup of a database, and that in order to access and query it, I would need to restore the database using SQL Server Management Studio. Once this was completed, I added my local SQL server and the restored database as an ODBC Data Source. Finally, I used the `DBI` library in R, in order to establish a connection to the database.


```{r}
# Libraries
library("DBI")
library("tidyverse")
```


```{r}
# Establishing ODBC connection to 'SQLserveR'
con <- dbConnect(odbc::odbc(), "SQLserveR")
```


## Toronto Blue Jays - Historic Regular Season Pitching Data

I am interested in pitching statistics for my favorite team, the Toronto Blue Jays. Using the query below, I have created a data frame that contains the regular season pitching data for every pitcher, in every year that the Jays have existed, minus this past season (1977-2023.) Included in this dataframe are the pitcher's first and last names from the `People` table of lahman2023, and a bevy of pitching statistics from the `Pitching` table.

```{sql connection=con, output.var = pitching_data}
SELECT Pitching.*, People.nameFirst, People.nameLast, People.throws
FROM lahman2023..Pitching
JOIN lahman2023..People
ON Pitching.playerID = People.playerID
WHERE Pitching.teamID = 'TOR'
ORDER BY Pitching.yearID
```


Using this dataframe, I wanted to look at two statistics in particular - Completed Games (CG) and Saves (SV). 

```{r}
# R Code
# Look at a few stats, including CG - Completed Game and examine how that's change historically, and SV - Saves to see great relievers
# Summarize by year, CG. Summarize by year, SV

p1 <- pitching_data %>% group_by(yearID) %>% summarize("Completed Games" = sum(CG), "Saves" = sum(SV))

p1_long <- p1 %>% gather("Statistic", "Count", -yearID)

ggplot(p1_long, aes(x = yearID, y = Count, fill = Statistic)) + geom_col(position = "dodge") + labs(title = "Pitching: Total Completed Games and Saves by Year\nToronto Blue Jays (1977-2023)", x = "Year")
```

Plotting (...)


## Toronto Blue Jays - Historic Post Season Pitching and Hitting Data

The Blue Jays won the World series in the year I was born, 1992, and the following year, 1993. Given I missed watching these seasons and post-seasons, I figure it would be interesting to take a look at the best performers of these postseasons. 

```{sql connection=con, output.var = postpitching_data}

```


```{sql connection=con, output.var = posthitting_data}

```


```{r}
# R Code
# Visualize starting pitchers and their ERAs


```


```{r}
# R Code
# Visualize hitter stats
```



## Salaries for the Championship Teams

ABCDE


```{sql connection=con, output.var = salary_data}

```


```{r}
# Compare salary situations for the 1992 and 1993 teams, composition spent on hitter/pitchers
```



## Conclusion